---
title: "FDA Audit Analysis"
author: "Nick Cruickshank"
date: "8/15/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load libraries, message=FALSE, warning=FALSE}
library(forcats)
library(lubridate)
library(qdapRegex)
library(readr)
library(readxl)
library(tidyverse)
library(scales)
library(stringr)
```

```{r load data}
classifications <- read_excel("Inspection_Classification_(10-1-2008_through_7-22-2020).xlsx", 
    col_types = c("text", "text", "text", 
        "text", "text", "text", "date", "text", 
        "text", "text"))

citations <- read_excel("Inspection_Citation_(10-1-2008_through_7-22-2020)_0.xlsx", 
    col_types = c("text", "text", "text", 
        "text", "date", "text", "text", "text", 
        "text"))

drug_obs <- read_csv("fda_drug_observation.csv", 
    col_types = cols(`Cite Id` = col_character()))

list_503b <- read_csv("list_503b.csv", col_types = cols(current_outsourcing_registration_date = col_date(format = "%m/%d/%Y"), 
    date_last_inspection = col_date(format = "%m/%d/%Y"), 
    initial_outsourcing_registration_date = col_date(format = "%m/%d/%Y")))
```

# Description of Datasets

## Classifications

A dataset of all project areas inspected by the FDA between 10-01-2008 and 07-22-2020 categorized by classification.

```{r}
classifications %>%
  head(10)
```


Source: https://www.fda.gov/inspections-compliance-enforcement-and-criminal-investigations/inspection-classification-database
Info on Columns: https://www.accessdata.fda.gov/scripts/inspsearch/searchfields.cfm
Key for Districts: https://www.fda.gov/inspections-compliance-enforcement-and-criminal-investigations/compliance-actions-and-activities/district-names-and-abbreviations


No. | Column Name         | Class | Description
----|---------------------|-------|-------------
01  | District            | chr   | District where company resides
02  | Legal Name          | chr   | Name of the company inspected.
03  | City                | chr   | City the company is located in.
04  | State               | chr   | Abbreviation for the state the company is located in
05  | Zip                 | chr   | Zip code for company address
06  | County/Area         | chr   | State the company is based in.
07  | Inspection End Date | date  | Date the inspection was concluded
08  | Center              | chr   | Sub-department of the FDA.
09  | Project Area        | chr   | Categorized by corresponding center.
10  | Classification      | chr   | Inspection classification with regards to compliance status

Centers include the following: 
 - CFSAN: Center for Food Safety and Applied Nutrition
 - CBER: Center for Biologics Evaluation and Research
 - CDER: Center for Drug Evaluation and Radiological Health
 - CVM: Center for Veterinary Medicine
 - CDRH: Office of Regulatory Affairs
 - CTP: Center for Tobacco Product(s)

Classifications include the following:
 - NAI: No Action Indicated
 - VAI: Voluntary Action Indicated
 - OAI: Official Action Indicated

## Citations

Source: https://www.fda.gov/inspections-compliance-enforcement-and-criminal-investigations/inspection-references/inspection-citation

A data set of all citations given by the FDA during audits generated from FDA Form 483 between 10-01-2008 and 07-22-2008.

```{r}
citations %>%
  head(10)
```

## List of 503B Outsourcing Pharmaceutical Companies

Source: https://www.fda.gov/drugs/human-drug-compounding/registered-outsourcing-facilities

```{r}
list_503b %>%
  head()
```

## Drug Observations

A section of the Inspections Observations data sheet strictly related to Drug Products. Lists the frequency of all given observations.

```{r}
drug_obs %>%
  head(10)
```

Source: https://www.fda.gov/inspections-compliance-enforcement-and-criminal-investigations/inspection-references/inspection-observations
More Info: https://www.fda.gov/inspections-compliance-enforcement-and-criminal-investigations/inspection-references/inspectional-observations-and-citations

No. | Column Name           | Class | Description
----|-----------------------|-------|-------------
01  | Citation Program Area | chr   | 
02  | Cire Id               | chr   | 
03  | Reference Number      | chr   | 
04  | Short Description     | chr   | 
05  | Long Description      | chr   | 
06  | Frequency             | dbl   | 

# Exploratory Analysis

What proportion of FDA Inspections in 2019 were Center for Drug Evaluation and Radiological Health?

```{r}
blank_theme <- theme_minimal() + 
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.border = element_blank(),
    panel.grid = element_blank(),
    axis.ticks = element_blank(),
    plot.title = element_text(size = 14, face = "bold")
  )

classifications %>%
  group_by(Center) %>%
  dplyr::summarise(
    Inspections = n()
  ) %>%
  ggplot() + 
  geom_bar(aes(x = "", y = Inspections, fill = Center), stat = "identity") + 
  scale_fill_manual(values = c(
    "CBER" = "grey95",
    "CDER" = "firebrick",
    "CDRH" = "grey85",
    "CFSAN" = "grey75",
    "CTP" = "grey65",
    "CVM" = "grey55"
  )) + 
  coord_polar("y", start = 0) + 
  theme_minimal() + 
  blank_theme +
  theme(axis.text.x = element_blank())
```

Of the CDER Inspections, what was the distribution of Inspection Classifications?

What is the distribution of inspections per capita by state? Make a heatmap of America.

```{r}
classifications %>%
  filter(Center == "CDER",
         `Project Area` == "Drug Quality Assurance",
         !is.na(State)) %>%
  group_by(State) %>%
  dplyr::summarise(
    Inspections = n()
  )
```

What is the distribution of Classifications by State?

What months of the year is the agency most likely to perform an inspection? Make a bar graph with fill = classification. Does this distribution look different in the region which Vermont belongs to?

```{r}
classifications %>%
  filter(Center == "CDER",
         `Project Area` == "Drug Quality Assurance",
         !is.na(State),
         District == "NWE") %>%
  separate(`Inspection End Date`, into = c("year", "month", "day"), sep = "-", remove = FALSE) %>%
  group_by(month) %>%
  dplyr::summarise(
    count = n(),
    NAI = sum(Classification == "NAI"),
    VAI = sum(Classification == "VAI"),
    OAI = sum(Classification == "OAI"),
  ) %>%
  mutate(month_abr = case_when(
    month == "01" ~ "Jan",
    month == "02" ~ "Feb",
    month == "03" ~ "Mar",
    month == "04" ~ "Apr",
    month == "05" ~ "May",
    month == "06" ~ "Jun",
    month == "07" ~ "Jul",
    month == "08" ~ "Aug",
    month == "09" ~ "Sep",
    month == "10" ~ "Oct",
    month == "11" ~ "Nov",
    month == "12" ~ "Dec"
  )) %>%
  pivot_longer(cols = c(NAI, VAI, OAI), names_to = "Classification", values_to = "Inspections") %>%
  mutate(class_sev = case_when(
    Classification == "NAI" ~ 1,
    Classification == "VAI" ~ 2,
    Classification == "OAI" ~ 3)) %>%
  rename(c(
    "m1" = "month",
    "Month" = "month_abr"
  )) %>%
  ggplot(aes(fct_reorder(Month, m1), Inspections)) + 
  geom_bar(aes(fill = fct_reorder(Classification, desc(class_sev))), stat = "identity") + 
  scale_fill_brewer(palette = "YlOrRd", direction = -1) + 
  labs(
    title = "FDA Audits in the New England Area are less frequent in the summer",
    subtitle = "Classification of 'Offical Action Indicated' most common in March and August",
    x = "Month",
    y = "Number of Inspections",
    fill = "Classification",
    caption = "Source: fda.gov"
  ) + 
  theme_bw() + 
  theme(
    legend.position = "bottom"
  )
```

Where are most drug inspections performed? Which state has the most drug companies? Make a heatmap of drug manufacturer's per capita.

## Citation Exploration

Merge Citations and List of 503B Companies

Problems to address:
 - Citations has a company named "ImprimisRx NJ" while tidy_503B has a company named "Imprimis NJOF". They need to match.
 - Citations has a company named "Pharmaceutics Labs" while tidy_503B has a company named "Pharamceutic Lab" (no 's' at the end). They need to match.

```{r tidy data, message=FALSE, warning=FALSE}
# tidy list_503b
tidy_503b <- list_503b %>%
  mutate(facility = gsub("[,]? (LLC|Inc)[\\.]?[,]?", ",", `Facility Name`)) %>%
  mutate(facility = gsub("[,]? (formerly|dba|DBA|d/b/a|dba:) ", " formerly ", facility)) %>%
  mutate(facility = gsub("355 Riverwalk Pkwy, ", "", facility)) %>%
  mutate(facility = gsub(", ((NSCC)), ", "", facility)) %>%
  separate(facility, into = c("name", "city", "state"), sep = ", ", remove = FALSE) %>%
  filter(!is.na(state),
         city != "(NSCC)",
         name != "STERRX") %>%
  separate(name, into = c("Current Name", "Formerly Name"), sep = " formerly ", remove = TRUE) %>%
  mutate(`Formerly Name` = gsub("registered as ", "", `Formerly Name`)) %>%
  separate(`Formerly Name`, into = c("Former Name", "c1", "s1"), sep = ", ", remove = FALSE) %>%
  select(`Current Name`, `Former Name`, city, state, initial_outsourcing_registration_date, current_outsourcing_registration_date,
         date_last_inspection, form483_issued, other_action, intent_to_compound_from_bulk_drug_substance) %>%
  rename(c(
    #"Facility Name" = "Current Name",
    "City" = "city",
    "State" = "state"
  )) %>%
  pivot_longer(cols = c("Current Name", "Former Name"), names_to = "Name Status", values_to = "Facility Name")

tidy_503b <- tidy_503b[complete.cases(tidy_503b), ]
tidy_503b$`Compounding Outsourcing Facility` = "Yes"

cite_503b <- citations %>%
  filter(`Program Area` == "Drugs") %>%
  mutate(fac = gsub("[,]? (LLC|Inc\\.|Lp|Ltd|Llc|LTD|LP|PLC|)[\\.]?", " ", `Firm Name`)) %>% # remove LLC or Inc from Firm Name
  mutate(fac = trimws(fac, which = "right")) %>%
  separate(fac, into = c("Current Name", "Former Name"), sep = "\\s(DBA|dba|d/b/a)\\s", remove = FALSE) %>%
  mutate(`Current Name` = trimws(`Current Name`, which = "right")) %>%
  pivot_longer(cols = c("Current Name", "Former Name"), names_to = "Name Status", values_to = "Facility Name") %>%
  merge(tidy_503b, by = c("Facility Name", "City", "State")) %>%
  select(`Facility Name`, City, State, `Country/Area`, `Inspection End Date`, `Program Area`,
         `CFR/Act Number`, `Short Description`, `Long Description`, `Name Status.x`,
         initial_outsourcing_registration_date, current_outsourcing_registration_date, date_last_inspection,
         form483_issued, other_action, intent_to_compound_from_bulk_drug_substance)

true_503B <- citations %>%
  filter(`Program Area` == "Drugs",
         `Country/Area` == "United States") %>%
  mutate(fac = gsub("[,]? (LLC|Inc\\.|Lp|Ltd|Llc|LTD|LP|PLC|)[\\.]?", " ", `Firm Name`)) %>% # remove LLC or Inc from Firm Name
  mutate(fac = trimws(fac, which = "right")) %>%
  separate(fac, into = c("Current Name", "Former Name"), sep = "\\s(DBA|dba|d/b/a)\\s", remove = FALSE) %>%
  mutate(`Current Name` = trimws(`Current Name`, which = "right")) %>%
  pivot_longer(cols = c("Current Name", "Former Name"), names_to = "Name Status", values_to = "Facility Name") %>%
  semi_join(tidy_503b, by = c("Facility Name", "City", "State")) %>%
  select(`Facility Name`, City, State, `Country/Area`, `Inspection End Date`, `Program Area`,
         `CFR/Act Number`, `Short Description`, `Long Description`)

true_503B$`Compounding Outsourcing Facility` = 1

false_503B <- citations %>%
  filter(`Program Area` == "Drugs",
         `Country/Area` == "United States") %>%
  mutate(fac = gsub("[,]? (LLC|Inc\\.|Lp|Ltd|Llc|LTD|LP|PLC|)[\\.]?", " ", `Firm Name`)) %>% # remove LLC or Inc from Firm Name
  mutate(fac = trimws(fac, which = "right")) %>%
  separate(fac, into = c("Current Name", "Former Name"), sep = "\\s(DBA|dba|d/b/a)\\s", remove = FALSE) %>%
  mutate(`Current Name` = trimws(`Current Name`, which = "right")) %>%
  pivot_longer(cols = c("Current Name", "Former Name"), names_to = "Name Status", values_to = "Facility Name") %>%
  anti_join(tidy_503b, by = c("Facility Name", "City", "State")) %>%
  select(`Facility Name`, City, State, `Country/Area`, `Inspection End Date`, `Program Area`,
         `CFR/Act Number`, `Short Description`, `Long Description`)

false_503B <- false_503B[complete.cases(false_503B), ]
false_503B$`Compounding Outsourcing Facility` = 1

tidy_citations <- rbind(true_503B, false_503B)
```

Can you merge citations and classifications?

```{r}
classifications %>%
  left_join(citations)
```

